Gusto: a toolkit for compatible finite element dynamical cores

### 2024-08-19 source/paper.tex
work on the intro

If the different finite element function spaces are chosen so that
discrete versions of vector calculus identities hold then the
discretisation is called `mimetic' or `compatible'.

* These methods are useful in a variety of applications
* In the context of geophysical fluid dynamics they are useful because...
* These properties rely on the underpinning structure of the de Rham complex
* In 3D the function spaces have a tensor product structure
* The function spaces are hard to construct especially on the sphere - piola transform

>>>>

If the different finite element function spaces are chosen so that
discrete versions of vector calculus identities hold then the
discretisation is called `mimetic' or `compatible'. These methods are
useful in a variety of applications due to their ability to capture
the underlying geometric structure of the governing equations. In the
context of geophysical fluid dynamics, these methods are useful
because they represent geostrophic balance exactly which is crucial
for reproducing accurate simulations of the large-scale flow. In
addition to this, there is flexibility within the compatible finite
element framework to choose function spaces that also give other
advantages such as energy conservation, local mass conservation and
discrete dispersion properties that lead to no spurious pressure
modes.

The compatible finite element spaces relevant for geophysical fluid
dynamics are constructed from a discrete differential complex called
the de Rham complex. This is a sequence of spaces with differential
operators between them that map onto the kernal of the next operator
in the sequence. This is often summarised in a commutation diagram as
in figures \ref{fig:3D de Rham} and \ref{fig:2D de Rham}. In these
diagrams the infinite dimensional spaces are along the top row and the
$\pi_i$ denote projections into the corresponding finite dimensional
subspace used in the compatible finite element discretisation. The
differential operators are denoted by $\diff^k$ and these commute with
the projections, meaning that $\diff^{k+1}\pi_k(\cdot) =
\diff^{k+1}\pi_{k+1}(\cdot)$. This guarantees that, in 3D, $\nabla$
maps onto the kernal of $\nabla\times$, ensuring
$\nabla\times\nabla(\cdot)=0$, and $\nabla\times$ maps onto the kernal
of $\nabla\cdot$, ensuring that $\nabla\cdot\nabla\times(\cdot)=0$;
i.e. the spaces are compatible. In 2D we have that $\nabla^\perp$ maps
onto the kernal of $\nabla\cdot$, ensuring that
$\nabla\cdot\nabla^\perp(\cdot)=0$. It is possible to write this de
Rham complex such that instead we have that $\nabla$ maps onto the
kernal of $\nabla^\perp$, ensuring that
$\nabla^\perp\cdot\nabla(\cdot)=0$, but the former is more useful in
constructing discretisations for geophysical fluid dynamics since it
enables schemes with local mass conservation \citep{}.

* In 3D the function spaces have a tensor product structure
* The function spaces are hard to construct especially on the sphere - piola transform, theta space
